A sheaf--function dictionary for algebraic tori over local fields

### 2013-09-23 shanghai.tex
More = -> \ceq

@@

\begin{lemma}\label{lemma:section}
 If $\qcs{L}$ is a quasicharacter sheaf on a commutative, \'etale group scheme $G$
 over $\Fq$ and if $G(\bFq)$ is finitely generated, then there is a global section
 $s = \{ s(x) \in \gqcs{L}_x \tq x \in G(\bFq) \}$ satisfying
 \begin{equation}\label{section}
 \end{equation}
\end{lemma}

@@

\begin{proof}
  Suppose $\qcs{L}$ is a quasicharacter sheaf on $G$.  
  Using Lemma~\ref{lemma:section}, 
  we may choose a section $s = \{ s(x) \in \gqcs{L}_x \tq x \in G(\bFq) \}$ satisfying \eqref{section}.
  We will construct a group isomorphism between $\QCiso{G}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ using $s$, then prove independence from the choice of $s$.

@@

  Although we used a global section $s$ in the definition of
  $\tau_\qcs{L}$, the class of $\tau_\qcs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  $\qcs{L}$ and let $\tau_1$ and $\tau_2$ be the cocycles defined by $s_1$ and $s_2$.
  For each $x \in G(\bFq)$ there is a unique scalar
  $a(x) \in \EEx$ such that $s_2(x) = a(x) s_1(x)$. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$.  The coboundary $\tau_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $\tau_0(\Frob{\Fq}) = \,^{\Frob{\Fq}} a \cdot a^{-1}$ satisfies $\tau_2 = \tau_0 \tau_1$,
  so $\tau_2$ and $\tau_1$ are cohomologous. Thus
  \begin{equation}\label{cohomologous}
    \obj \QC(G) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \qcs{L} \mapsto [\tau_\qcs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

@@

 To see that \eqref{pich1} is injective,
 suppose $[\tau_\qcs{L}] = [\tau_\qcs{L'}]$.
 Let $s$ (resp. $s'$) be a section appearing in the definition of $\tau_\qcs{L}$ (resp. $\tau_\qcs{L'}$);
 recall that $s$ (resp. $s'$) is compatible with $\mu$ (resp $\mu'$) 
 in the sense of Lemma~\ref{lemma:section}.
 Since $[\tau_\qcs{L}] = [\tau_\qcs{L'}]$, 
 it follows that $\tau_\qcs{L}$ and $\tau_{\qcs{L'}}$ are cohomologous, 
  so there is some $a\in G(\Fq)^*$ such that 
  $\tau'(\Frob{\Fq}) = \tau(\Frob{\Fq}) \cdot (\,^{\Frob{\Fq}}a \cdot a^{-1})$. 
  Define $\alpha : \qcs{L}\to \qcs{L'}$ by $\alpha_x(s(x)) = a(x) s'(x)$, for each $x\in G(\bFq)$.
  It is straightforward to check that $\alpha$ satisfies condition~\ref{cs.3}
  and that $\alpha^{-1} : \qcs{L'}\to \qcs{L}$, defined by $\alpha^{-1}_x(s'(x)) = a(x)^{-1} s(x)$, 
  is its inverse. Thus, $\qcs{L} \iso \qcs{L'}$ in $\QC(G)$. 
  This shows that \eqref{pich1} is injective.

 Finally, to show that \eqref{pich1} is surjective, pick $[\tau]\in \Hh^1(\Weil{\Fq}, G(\bFq)^*)$.
 For each $x\in G(\bFq)$, set $\gqcs{L}_x = \EE$.
 Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by
 $\mu_{x,y}(1) = 1\otimes1$, for $x,y\in G(\bFq)$; 
 clearly, $\mu$ satisfies condition~\ref{cs.1}.
 Define $\phi_x : \gqcs{L}_{\Frob{\Fq}} \to \gqcs{L}_x$ by $\phi_x(1) = \tau(\Frob{\Fq})(x)$,
 for each $x\in G(\bFq)$. 
 Then $\phi$ satisfies condition~\ref{cs.2} because $\tau \in Z^1(\Weil{\Fq},G(\bFq)^*)$.
 In this way we have built a quasicharacter sheaf $\qcs{L} = (\gqcs{L},\mu,\phi)$ 
 from $\tau \in \Hh^1(\Weil{\Fq}, G(\bFq)^*)$ such that the image of $\qcs{L}_{/\text{iso}}$
 under \eqref{pich1} is the class $[\tau]$ in $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$.
 This shows that \eqref{pich1} is surjective, 
 and completes the proof that it is an isomorphism of groups.
\end{proof}

@@

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
\[
    \Hh^1(\Fq, X^*) \xrightarrow{\rho} \Hh^0(\Fq, X)^*,
 \qquad \text{defined by}\qquad
   \rho([z])(x) = z(\Frob{\Fq})(x),
\]
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^{\Frob{\Fq}}, \EEx).$ Since $\EEx$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^{\Frob{\Fq}}, \EEx) = 0$.  Applying the functor
  $\Hom(\mbox{---}, \EEx)$ to
  \[
  \begin{tikzcd}
  0 \rar & X^{\Frob{\Fq}} \rar & X \rar & X/X^{\Frob{\Fq}} \rar & 0
  \end{tikzcd}
  \]
  and considering the associated long exact sequence gives the
  existence of an $\tilde{f} \in \Hom(X, \EEx)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EEx))$ by
  mapping $\Frob{\Fq}$ to $\tilde{f}$.  Then $\rho([z]) = f$.

@@

  We now fix a $d$, $P$ and $s$, write $\zeta$ for $\zeta_d$ and
  choose a polynomial
  $$h(T) = a_rT^r + a_{r-1}T^{r-1} + \cdots + a_0$$
  so that $P^s = (p^s, h(\zeta_d))$.
  % Setting $\psi = 1$ will satisfy Equation \ref{eq:psi-condition}
  % when $d = 1$.
  An an abelian group, the quotient $\ZZ[\zeta] / P^s$ is generated by
  $\zeta^i$ for $0 \le i < \phi(d)$; we define
  \begin{equation} \label{eq:psi-def}
    \psi(\zeta^i) = \alpha \prod_{j=0}^{i-1} \tilde{f}(\zeta^j)
  \end{equation}
  for an $\alpha \in \EEx$ to be chosen later.  Since there are
  relations among the $\zeta^i$ we first need to check that $\psi$ is
  well-defined.

  The relations among the generators $\zeta^i$ are additively
  generated by $\zeta^ch(\zeta)$ as $c$ ranges from $0$ to
  $\phi(d)-r-1$.  If we set
  $$A_c = \sum_{i=0}^r a_i \sum_{j=0}^{i+c-1} \zeta^j$$
  then
  $$\psi(\zeta^ch(\zeta)) = \alpha^{h(1)} \tilde{f}(A_c).$$
  Now consider $(\zeta - 1)A_c$: the sums telescope and we get that
  $(\zeta-1)A_c = \zeta^ch(\zeta) - h(1).$ There are now two
  possibilities:
  \begin{enumerate}
  \item $d$ is a power of $p$; write $s = \phi(d)m + n$.  We may take
    $h(T)$ to be $p^m(T-1)^n$, in which case $\zeta^ch(\zeta) - h(1) \in P^s$,
    so $A_c$ is fixed under multiplication by $\zeta$ in
    $\ZZ[\zeta]/P^s$ and thus $\tilde{f}(A_c) = 1$.  In this case
    $\alpha^{h(1)} = 1$ regardless of $\alpha$.
  \item $d$ is not a power of $p$.  If two distinct primes divide $d$
    then $\zeta - 1$ is a unit; otherwise $\zeta - 1$ is a unit modulo
    $P^s$.  Let $\omega \in \ZZ[\zeta]$ satisfy
    $\omega(\zeta - 1) \equiv 1 \pmod{P^s}$ and set
    %\begin{equation} \label{alpha-def}
     $ \alpha = \tilde{f}(\omega)$.
    %\end{equation}
    Then $A_c + h(1)\omega$ is fixed under multiplication by $\zeta$
    and thus $\tilde{f}(A_c) = \tilde{f}(-h(1)\omega)$.
  \end{enumerate}
  In either case, $\psi(\zeta^ch(\zeta)) = 1$ for all $c$.  The only
  other relations imposed on the $\zeta^i$ are that they are all
  killed under multiplication by $p^s$.  Since $\tilde{f}$ is also a
  homomorphism $\ZZ[\zeta]/P^s \rightarrow \EEx$ the values of
  $\tilde{f}$ and thus of $\psi$ will be $p^s$-roots of unity.
  Therefore $\psi$ is well defined.

  We now check Equation~\eqref{eq:psi-condition} for
  $x = \zeta^{\phi(d)-1}$.  Write
  \[\Phi_d(T) = T^{\phi(d)} - b_{\phi(d)-1}T^{\phi(d)-1} - \cdots - b_0\]
  and set
  $$B = b_i \sum_{j=0}^{i-1} \zeta^j -  \sum_{i=0}^{\phi(d)-1} \zeta^i.$$
  We have

@@

  \begin{enumerate}
  \item $d$ is a power of $p$, in which case $\Phi_d(1) = p$.  Thus
    $(\zeta-1)B = p = u(\zeta-1)^{\phi(d)}$ for a unit $u \in \ZZ[\zeta]$
    and $B - u(\zeta-1)^{\phi(d)-1}$ is fixed by $\zeta$,
    so that $\tilde{f}(B) = \tilde{f}(u(\zeta - 1)^{\phi(d)-1})$.  The
    multiplicative order of $\tilde{f}(u(\zeta-1)^{\phi(d)-1})$ is at most
    $\lceil \frac{s - 1 - (\phi(d)-1)}{\phi(d)}\rceil = \lceil \frac{s}{\phi(d)} \rceil - 1$
    while the additive order of $1 \in \ZZ[\zeta]/P^s$ is $\lceil \frac{s}{\phi(d)} \rceil$.
    Thus we may choose $\alpha = \psi(1)$ with $\alpha^p = \tilde{f}(B)$.
  \item $d$ is not a power of $p$ and $\omega$ is the inverse of
    $\zeta-1$ modulo $P^s$.  Then $\tilde{f}(B) = \tilde{f}(\omega\Phi_d(1))$.
  \end{enumerate}
  In either case, $\alpha^{-\Phi_d(1)}\tilde{f}(B) = 1$ and thus
  $\psi$ satisfies Equation~\eqref{eq:psi-condition} for
  $x = \zeta^{\phi(d)-1}$.  The definition of $\psi$ in \ref{eq:psi-def}
  immediately implies that $\psi$ satisfies equation
  \ref{eq:psi-condition} for lesser powers of $\zeta$, and thus for
  all $x$ by linearity.  The existence of $\psi$ shows that $\rho$ is injective.
\end{proof}

@@

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny; set $A = \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G)
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

  Set $\Gamma = \Gal(\bFq/\Fq)$. Extensions of
  $G^0(\bFq)$ by $A(\bFq)$, such as $A(\bFq) \to H^0(\bFq) \to G^0(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$,
  fitting into the long exact sequence derived from applying
  the functor $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  \[
  \Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)).
  \]
  We now show that $\Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)) = 0$.

  It suffices to show that $\pi_0(G)(\bFq)$ has a two-term resolution
  by injective $\ZZ[\Gamma]$-modules.  Since the action of $\Gamma$ on
  $\pi_0(G)(\bFq)$ is continuous, it factor through a finite quotient
  $\Gamma_m = \Gamma / H_m$.  Any product of Dedekind domains has
  cohomological dimension $1$ \David{Add a citation for this fact}, so $\pi_0(G)(\bFq)$ has a
  two term resolution by injective $\ZZ[\Gamma_m]$-modules.  Let $I$
  be one of these modules: we need to show that $I$ is injective as a
  $\ZZ[\Gamma]$-module.  Consider the universal diagram for injectivity:

@@

A N\'eron model for $T$, denoted here by $\TT$ 
and whose existence is given by \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch 10.1, Thm 6},
is a separated, smooth, commutative group scheme locally of finite type over $\OK$ with generic fibre $T$,
that enjoys the N\'eron mapping property:
if $S$ is a smooth scheme over $\OK$ then every morphism $S_\eta \to T$
over $K$ extends uniquely to a morphism $S \to \TT$. 
Equivalently, the
N\'eron mapping property states that the function
$\Hom_{\OK} (S,\TT) \to \Hom_K(S_\eta,T)$, given by restriction to
generic fibres, is a bijection; 
in particular, taking $S = \Spec{\OK}$, this implies $\TT(\OK) = T(K)$.

@@

\begin{definition}\label{def:Grn}
For any local field $K$ with residue field $\Fq$ and for any positive integer $n$, let
\[
%\begin{tikzcd}
\Gr_n : \left( \text{Schemes} \atop \text{lft over $\OK$}\right) \to \left( \text{Schemes} \atop \text{lft over $\Fq$}\right)
%\end{tikzcd}
\]
be the functor produced by precomposing pull-back along $\OK \to \OK/\pK^n$ with the Greenberg functor.
\end{definition}

@@

\begin{remark}
If $K = \mathbb{Q}_q$ and 
$T = \Gm{\mathbb{Q}_q}$ then $\GN{T}^0 = \mathbb{W}^*_{\FF_q}$, 
the group scheme (!) over $\Fq$ of units in the ring scheme $\mathbb{W}_{\FF_q}$
and $\GN{T}$, as a scheme, is the disjoint union of copies of $\mathbb{W}^*_{\FF_q}$
indexed by the integers. Based on this example, 
we ask if, for arbitrary $T$, the pro-group scheme $\GN{T}$ is, in fact, a group scheme.
We have not explored that question.
\end{remark}

@@

\begin{proposition}\label{prop:wrK}
Let $K'/K$ be a finite Galois extension and
let $k'/k$ be the corresponding finite Galois extension of residue fields.
Let $T$ be an algebraic torus over $K$ and 
set $T' = T \times_\Spec{K} \Spec{K'}$.
The canonical closed immersion $T \hookrightarrow \Res_{K'/K} T'$
induces a map of group pro-schemes $\GN{T} \to \Res_{k'/k} \GN{T}'$ which,
through the pull-back functor and the trace of Frobenius, induces the homomorphism
\[
\begin{tikzcd}[column sep=60]
\Hom_{\text{qc}}(T(K'), \EEx) \arrow{r}{\chi \mapsto \chi\vert_{T(K)}} &\Hom_{\text{qc}}(T(K), \EEx).
\end{tikzcd}
\]  
\end{proposition}

>>>>

%amsmath replacements for \atop
\newcommand{\latop}[2]{\genfrac{}{}{0pt}{0}{#1}{#2}}
\newcommand{\partop}[2]{\genfrac{(}{)}{0pt}{0}{#1}{#2}}

@@

\begin{lemma}\label{lemma:section}
 If $\qcs{L}$ is a quasicharacter sheaf on a commutative, \'etale group scheme $G$
 over $\Fq$ and if $G(\bFq)$ is finitely generated, then there is a global section
 $s \ceq \{ s(x) \in \gqcs{L}_x \tq x \in G(\bFq) \}$ satisfying
 \begin{equation}\label{section}
 \end{equation}
\end{lemma}

@@

\begin{proof}
  Suppose $\qcs{L}$ is a quasicharacter sheaf on $G$.  
  Using Lemma~\ref{lemma:section}, 
  we may choose a section $s \ceq \{ s(x) \in \gqcs{L}_x \tq x \in G(\bFq) \}$ satisfying \eqref{section}.
  We will construct a group isomorphism between $\QCiso{G}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ using $s$, then prove independence from the choice of $s$.

@@

  Although we used a global section $s$ in the definition of
  $\tau_\qcs{L}$, the class of $\tau_\qcs{L}$ does not depend on this
  choice. To see why, let $s_1$ and $s_2$ be two global sections of
  $\qcs{L}$ and let $\tau_1$ and $\tau_2$ be the cocycles defined by $s_1$ and $s_2$.
  For each $x \in G(\bFq)$ there is a unique scalar
  $a(x) \in \EEx$ such that $s_2(x) = a(x) s_1(x)$. Then \eqref{section} and
  condition~\ref{cs.1} together imply $a(x+y) = a(x)a(y)$, so
  $a\in G(\bFq)^*$.  The coboundary $\tau_0 : \Weil{\Fq} \to G(\bFq)^*$
  defined by $\tau_0(\Frob{\Fq}) \ceq \,^{\Frob{\Fq}} a \cdot a^{-1}$ satisfies $\tau_2 = \tau_0 \tau_1$,
  so $\tau_2$ and $\tau_1$ are cohomologous. Thus
  \begin{equation}\label{cohomologous}
    \obj \QC(G) \longrightarrow \Hh^1(\Weil{\Fq},G(\bFq)^*),
    \qquad \text{defined by} \qquad \qcs{L} \mapsto [\tau_\qcs{L}],
  \end{equation}
  is independent of the choice of global sections made above.

@@

 To see that \eqref{pich1} is injective,
 suppose $[\tau_\qcs{L}] = [\tau_\qcs{L'}]$.
 Let $s$ (resp. $s'$) be a section appearing in the definition of $\tau_\qcs{L}$ (resp. $\tau_\qcs{L'}$);
 recall that $s$ (resp. $s'$) is compatible with $\mu$ (resp $\mu'$) 
 in the sense of Lemma~\ref{lemma:section}.
 Since $[\tau_\qcs{L}] = [\tau_\qcs{L'}]$, 
 it follows that $\tau_\qcs{L}$ and $\tau_{\qcs{L'}}$ are cohomologous, 
  so there is some $a\in G(\Fq)^*$ such that 
  $\tau'(\Frob{\Fq}) = \tau(\Frob{\Fq}) \cdot (\,^{\Frob{\Fq}}a \cdot a^{-1})$. 
  Define $\alpha : \qcs{L}\to \qcs{L'}$ by $\alpha_x(s(x)) \ceq a(x) s'(x)$, for each $x\in G(\bFq)$.
  It is straightforward to check that $\alpha$ satisfies condition~\ref{cs.3}
  and that $\alpha^{-1} : \qcs{L'}\to \qcs{L}$, defined by $\alpha^{-1}_x(s'(x)) \ceq a(x)^{-1} s(x)$, 
  is its inverse. Thus, $\qcs{L} \iso \qcs{L'}$ in $\QC(G)$. 
  This shows that \eqref{pich1} is injective.

 Finally, to show that \eqref{pich1} is surjective, pick $[\tau]\in \Hh^1(\Weil{\Fq}, G(\bFq)^*)$.
 For each $x\in G(\bFq)$, set $\gqcs{L}_x \ceq \EE$.
 Define $\mu_{x,y} : \gqcs{L}_{x+y} \to \gqcs{L}_x\otimes \gqcs{L}_y$ by
 $\mu_{x,y}(1) \ceq 1\otimes1$, for $x,y\in G(\bFq)$; 
 clearly, $\mu$ satisfies condition~\ref{cs.1}.
 Define $\phi_x : \gqcs{L}_{\Frob{\Fq}} \to \gqcs{L}_x$ by $\phi_x(1) \ceq \tau(\Frob{\Fq})(x)$,
 for each $x\in G(\bFq)$. 
 Then $\phi$ satisfies condition~\ref{cs.2} because $\tau \in Z^1(\Weil{\Fq},G(\bFq)^*)$.
 In this way we have built a quasicharacter sheaf $\qcs{L} \ceq (\gqcs{L},\mu,\phi)$ 
 from $\tau \in \Hh^1(\Weil{\Fq}, G(\bFq)^*)$ such that the image of $\qcs{L}_{/\text{iso}}$
 under \eqref{pich1} is the class $[\tau]$ in $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$.
 This shows that \eqref{pich1} is surjective, 
 and completes the proof that it is an isomorphism of groups.
\end{proof}

@@

\begin{proposition}\label{prop:X}
  Let $X$ be a finitely generated abelian group equipped with an
  action of $\Gal(\bFq/\Fq)$, continuous for the discrete topology on
  $X$.  Then
\[
    \Hh^1(\Fq, X^*) \xrightarrow{\rho} \Hh^0(\Fq, X)^*,
 \qquad \text{defined by}\qquad
   \rho([z])(x) \ceq z(\Frob{\Fq})(x),
\]
  is an isomorphism of groups.
\end{proposition}
\begin{proof}
  We first prove that $\rho$ is surjective.  Suppose that
  $f \in \Hom(X^{\Frob{\Fq}}, \EEx).$ Since $\EEx$ is divisible
  as an abelian group it is injective as a $\ZZ$-module and thus
  $\Ext^1_{\ZZ}(X/X^{\Frob{\Fq}}, \EEx) = 0$.  Applying the functor
  $\Hom(\mbox{---}, \EEx)$ to
  \[
  \begin{tikzcd}
  0 \rar & X^{\Frob{\Fq}} \rar & X \rar & X/X^{\Frob{\Fq}} \rar & 0
  \end{tikzcd}
  \]
  and considering the associated long exact sequence gives the
  existence of an $\tilde{f} \in \Hom(X, \EEx)$ restricting to
  $f$.  Define a $1$-cocycle $z \in \Hh^1(\Fq, \Hom(X,\EEx))$ by
  mapping $\Frob{\Fq}$ to $\tilde{f}$.  Then $\rho([z]) = f$.

@@

  We now fix a $d$, $P$ and $s$, write $\zeta$ for $\zeta_d$ and
  choose a polynomial
  $$h(T) \ceq a_rT^r + a_{r-1}T^{r-1} + \cdots + a_0$$
  so that $P^s = (p^s, h(\zeta_d))$.
  % Setting $\psi = 1$ will satisfy Equation \ref{eq:psi-condition}
  % when $d = 1$.
  An an abelian group, the quotient $\ZZ[\zeta] / P^s$ is generated by
  $\zeta^i$ for $0 \le i < \phi(d)$; we define
  \begin{equation} \label{eq:psi-def}
    \psi(\zeta^i) \ceq \alpha \prod_{j=0}^{i-1} \tilde{f}(\zeta^j)
  \end{equation}
  for an $\alpha \in \EEx$ to be chosen later.  Since there are
  relations among the $\zeta^i$ we first need to check that $\psi$ is
  well-defined.

  The relations among the generators $\zeta^i$ are additively
  generated by $\zeta^ch(\zeta)$ as $c$ ranges from $0$ to
  $\phi(d)-r-1$.  If we set
  $$A_c \ceq \sum_{i=0}^r a_i \sum_{j=0}^{i+c-1} \zeta^j$$
  then
  $$\psi(\zeta^ch(\zeta)) = \alpha^{h(1)} \tilde{f}(A_c).$$
  Now consider $(\zeta - 1)A_c$: the sums telescope and we get that
  $(\zeta-1)A_c = \zeta^ch(\zeta) - h(1).$ There are now two
  possibilities:
  \begin{enumerate}
  \item $d$ is a power of $p$; write $s = \phi(d)m + n$.  We may take
    $h(T)$ to be $p^m(T-1)^n$, in which case $\zeta^ch(\zeta) - h(1) \in P^s$,
    so $A_c$ is fixed under multiplication by $\zeta$ in
    $\ZZ[\zeta]/P^s$ and thus $\tilde{f}(A_c) = 1$.  In this case
    $\alpha^{h(1)} = 1$ regardless of $\alpha$.
  \item $d$ is not a power of $p$.  If two distinct primes divide $d$
    then $\zeta - 1$ is a unit; otherwise $\zeta - 1$ is a unit modulo
    $P^s$.  Let $\omega \in \ZZ[\zeta]$ satisfy
    $\omega(\zeta - 1) \equiv 1 \pmod{P^s}$ and set
    %\begin{equation} \label{alpha-def}
     $ \alpha \ceq \tilde{f}(\omega)$.
    %\end{equation}
    Then $A_c + h(1)\omega$ is fixed under multiplication by $\zeta$
    and thus $\tilde{f}(A_c) = \tilde{f}(-h(1)\omega)$.
  \end{enumerate}
  In either case, $\psi(\zeta^ch(\zeta)) = 1$ for all $c$.  The only
  other relations imposed on the $\zeta^i$ are that they are all
  killed under multiplication by $p^s$.  Since $\tilde{f}$ is also a
  homomorphism $\ZZ[\zeta]/P^s \rightarrow \EEx$ the values of
  $\tilde{f}$ and thus of $\psi$ will be $p^s$-roots of unity.
  Therefore $\psi$ is well defined.

  We now check Equation~\eqref{eq:psi-condition} for
  $x \ceq \zeta^{\phi(d)-1}$.  Write
  \[\Phi_d(T) \ceq T^{\phi(d)} - b_{\phi(d)-1}T^{\phi(d)-1} - \cdots - b_0\]
  and set
  $$B \ceq b_i \sum_{j=0}^{i-1} \zeta^j -  \sum_{i=0}^{\phi(d)-1} \zeta^i.$$
  We have

@@

  \begin{enumerate}
  \item $d$ is a power of $p$, in which case $\Phi_d(1) = p$.  Thus
    $(\zeta-1)B = p = u(\zeta-1)^{\phi(d)}$ for a unit $u \in \ZZ[\zeta]$
    and $B - u(\zeta-1)^{\phi(d)-1}$ is fixed by $\zeta$,
    so that $\tilde{f}(B) = \tilde{f}(u(\zeta - 1)^{\phi(d)-1})$.  The
    multiplicative order of $\tilde{f}(u(\zeta-1)^{\phi(d)-1})$ is at most
    $\lceil \frac{s - 1 - (\phi(d)-1)}{\phi(d)}\rceil = \lceil \frac{s}{\phi(d)} \rceil - 1$
    while the additive order of $1 \in \ZZ[\zeta]/P^s$ is $\lceil \frac{s}{\phi(d)} \rceil$.
    Thus we may choose $\alpha = \psi(1)$ with $\alpha^p = \tilde{f}(B)$.
  \item $d$ is not a power of $p$ and $\omega$ is the inverse of
    $\zeta-1$ modulo $P^s$.  Then $\tilde{f}(B) = \tilde{f}(\omega\Phi_d(1))$.
  \end{enumerate}
  In either case, $\alpha^{-\Phi_d(1)}\tilde{f}(B) = 1$ and thus
  $\psi$ satisfies Equation~\eqref{eq:psi-condition} for
  $x \ceq \zeta^{\phi(d)-1}$.  The definition of $\psi$ in \ref{eq:psi-def}
  immediately implies that $\psi$ satisfies equation
  \ref{eq:psi-condition} for lesser powers of $\zeta$, and thus for
  all $x$ by linearity.  The existence of $\psi$ shows that $\rho$ is injective.
\end{proof}

@@

\begin{proof}
Let $\pi: B \to G^0$ be a discrete isogeny; set $A \ceq \ker \pi$.
  We will find a discrete isogeny $f: H\to G$
  such that that $H^0 = B$, $f^0 =\pi$ and
  $\pi_0(f) : \pi_0(H)\to \pi_0(G)$ is an isomorphism of component
  groups.
  \begin{equation}\label{extension-diagram}
  \begin{tikzcd}
  A \arrow[equal]{r} \dar & A \dar \\
  B \rar \dar[swap]{\pi} & H \rar \dar[swap]{f} & \pi_0(H) \arrow{d}[below,rotate=90]{\sim}[swap]{\pi_0(f)} \\
  G^0 \rar & G \rar & \pi_0(G)
  \end{tikzcd}
  \end{equation}
  where all rows and columns are exact and all maps are defined over
  $\Fq$.  We will do so by passing back and forth between group
  schemes over $\Fq$ and their $\bFq$-points.

  Set $\Gamma \ceq \Gal(\bFq/\Fq)$. Extensions of
  $G^0(\bFq)$ by $A(\bFq)$, such as $A(\bFq) \to H^0(\bFq) \to G^0(\bFq)$,
  correspond to classes in $\Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$.
  Similarly, extensions of $G(\bFq)$ by $A(\bFq)$ correspond to
  classes in $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq))$.  The map
  $G^0(\bFq) \to G(\bFq)$ induces the map
  $\Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq))$,
  fitting into the long exact sequence derived from applying
  the functor $\Hom(\mbox{---}, A(\bFq))$ to $G^0(\bFq) \to G(\bFq) \to \pi_0(G)(\bFq)$:
  \[
  \Ext^1_{\ZZ[\Gamma]}(G(\bFq), A(\bFq)) \to \Ext^1_{\ZZ[\Gamma]}(G^0(\bFq), A(\bFq)) \to \Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)).
  \]
  We now show that $\Ext^2_{\ZZ[\Gamma]}(\pi_0(G)(\bFq), A(\bFq)) = 0$.

  It suffices to show that $\pi_0(G)(\bFq)$ has a two-term resolution
  by injective $\ZZ[\Gamma]$-modules.  Since the action of $\Gamma$ on
  $\pi_0(G)(\bFq)$ is continuous, it factor through a finite quotient
  $\Gamma_m \ceq \Gamma / H_m$.  Any product of Dedekind domains has
  cohomological dimension $1$ \David{Add a citation for this fact}, so $\pi_0(G)(\bFq)$ has a
  two term resolution by injective $\ZZ[\Gamma_m]$-modules.  Let $I$
  be one of these modules: we need to show that $I$ is injective as a
  $\ZZ[\Gamma]$-module.  Consider the universal diagram for injectivity:

@@

A N\'eron model for $T$, denoted here by $\TT$ 
and whose existence is given by \cite{bosch-lutkebohmert-reynaud:NeronModels}*{Ch 10.1, Thm 6},
is a separated, smooth, commutative group scheme locally of finite type over $\OK$ with generic fibre $T$,
that enjoys the N\'eron mapping property:
if $S$ is a smooth scheme over $\OK$ then every morphism $S_\eta \to T$
over $K$ extends uniquely to a morphism $S \to \TT$. 
Equivalently, the
N\'eron mapping property states that the function
$\Hom_{\OK} (S,\TT) \to \Hom_K(S_\eta,T)$, given by restriction to
generic fibres, is a bijection; 
in particular, taking $S \ceq \Spec{\OK}$, this implies $\TT(\OK) = T(K)$.

@@

\begin{definition}\label{def:Grn}
For any local field $K$ with residue field $\Fq$ and for any positive integer $n$, let
\[
%\begin{tikzcd}
\Gr_n : \partop{\text{Schemes}}{\text{lft over $\OK$}} \to \partop{\text{Schemes}}{\text{lft over $\Fq$}}
%\end{tikzcd}
\]
be the functor produced by precomposing pull-back along $\OK \to \OK/\pK^n$ with the Greenberg functor.
\end{definition}

@@

\begin{remark}
If $K \ceq \mathbb{Q}_q$ and 
$T \ceq \Gm{\mathbb{Q}_q}$ then $\GN{T}^0 = \mathbb{W}^*_{\FF_q}$, 
the group scheme (!) over $\Fq$ of units in the ring scheme $\mathbb{W}_{\FF_q}$
and $\GN{T}$, as a scheme, is the disjoint union of copies of $\mathbb{W}^*_{\FF_q}$
indexed by the integers. Based on this example, 
we ask if, for arbitrary $T$, the pro-group scheme $\GN{T}$ is, in fact, a group scheme.
We have not explored that question.
\end{remark}

@@

\begin{proposition}\label{prop:wrK}
Let $K'/K$ be a finite Galois extension and
let $k'/k$ be the corresponding finite Galois extension of residue fields.
Let $T$ be an algebraic torus over $K$ and 
set $T' \ceq T \times_\Spec{K} \Spec{K'}$.
The canonical closed immersion $T \hookrightarrow \Res_{K'/K} T'$
induces a map of group pro-schemes $\GN{T} \to \Res_{k'/k} \GN{T}'$ which,
through the pull-back functor and the trace of Frobenius, induces the homomorphism
\[
\begin{tikzcd}[column sep=60]
\Hom_{\text{qc}}(T(K'), \EEx) \arrow{r}{\chi \mapsto \chi\vert_{T(K)}} &\Hom_{\text{qc}}(T(K), \EEx).
\end{tikzcd}
\]  
\end{proposition}
